---
title: "Regional Frequency Analysis: Sliding Window Approach"
author: "Carlos Dobler -"
date: "`r Sys.Date()`"
output: github_document

---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = F,
                      message = F,
                      warning = F,
                      out.width="95%",
                      fig.height=4.1*3,#32, 
                      fig.width=9.7)


library(tidyverse)
library(lubridate)
library(stars)
library(furrr)
library(units)
library(tmap)
library(colorspace)

options(future.fork.enable = T)
plan(multicore)

dir_tmp <- "/mnt/pers_disk"

```

The following figure shows the distributions of regions by size. The most common size is of between ~30,000 and ~260,000 km^2. The distribution is heavily left-tailed skewed (small regions are way more common than very large ones; the skew is not immediately noticeable due to the x-axis scaling).

```{r, fig.width=6.6, fig.height=4, out.width="65%"}

reg <- 
  readRDS(here::here("out", "l_sk_wl.rds")) %>% 
  pluck(1)

reg <- 
  reg %>% 
  mutate(area = st_area(reg),
         area = set_units(area, km^2)) %>% 
  drop_units()


ggplot(reg, aes(area)) +
  geom_histogram() +
  scale_x_continuous(trans = "log2", 
                     breaks = scales::trans_breaks("log2", function(x) round(2^x, -2), n = 10),
                     labels = scales::comma) +
  labs(x = "km^2",
       caption = "Note: x-axis is log2-scaled")


```

---

In the next figure, the top map shows the extreme precipitation deciles at 0.5 degrees of warming. Places in blue (yellow) represent where the 100-yr storm precipitation amounts are the highest (lowest) in the world under the baseline. The middle and the bottom map show deciles of positive _change_ in extreme precipitation compared to the baseline (i.e. only increases in amounts). Places in blue (yellow) represent where 100-yr storms will intensify the most (least). Places in grey correspond to places that will experience a negative change (i.e. the 100-yr storm precipitation will decrease).



```{r, fig.height=4.8*3}

s <- 
  "/mnt/pers_disk/lev_and_prob_mosaics/mos_99perc_a1_staticregs.nc" %>% 
  read_ncdf() 

land <- 
  "/mnt/bucket_mine/misc_data/ne_50m_land/ne_50m_land.shp" %>% 
  st_read(quiet = T) %>% 
  mutate(a = 1) %>% 
  select(a) %>% 
  st_rasterize(s %>% slice(wl, 1) %>% mutate(a = NA) %>% select(a))

s[is.na(land)] <- NA

l_p <- 
  imap(seq(0.5, 3, by = 0.5) %>% as.character, function(wl, wl_i){
    
    ss <- 
      s %>% 
      slice(wl, wl_i) %>% 
      setNames("v") %>% 
      mutate(v = round(v, 1))
    
    d <- 
      ss %>% 
      pull() %>% 
      .[. > 0] %>%
      quantile(seq(0,1, by = 0.1), na.rm = T)
    
    tb_neg <- 
      ss %>% 
      as_tibble() %>% 
      mutate(wl = {{wl}}) %>% 
      filter(v < 0) %>% 
      mutate(v = factor(-1))
    
    tb <-
      ss %>% 
      mutate(v = ifelse(v < 0, NA, v)) %>% 
      cut(d) %>% 
      # mutate(v = as.integer(v)) %>%
      as_tibble() %>% 
      mutate(wl = {{wl}}) %>% 
      bind_rows(tb_neg)
    
    ggplot() +
      geom_raster(data = tb %>% filter(v != "-1"), aes(lon, lat, fill = v)) +
      geom_raster(data = tb %>% filter(v == "-1"), aes(lon, lat), fill = "grey70") +
      coord_equal(ylim = c(-60, 78), xlim = c(-160, 155)) +
      theme(axis.title = element_blank(),
            legend.position = "bottom") +
      
      scale_x_continuous(expand = c(0, 0)) +
      scale_y_continuous(expand = c(0, 0)) +
      
      guides(fill = guide_legend(nrow = 2, keywidth = 0.8, keyheight = 0.8, byrow = T)) +
      
      scale_fill_discrete_sequential("plasma",
                                     rev = T,
                                     na.value = "transparent",
                                     name = str_glue("Decile   ")) +
      labs(subtitle = str_glue("WL: {wl} degrees"))
    
  })


patchwork::wrap_plots(l_p[c(1,3,6)], ncol = 1)



```

--- 

The following figure shows positive _changes_ in the frequency of the 100-yr storm for varying degrees of warming relative to a 0.5 degree warming level. Places in blue (yellow) represent where the 100-yr storm frequency is expected to increase the most (least). Places in grey correspond to places that will experience a negative change (i.e. the 0.5 WL 100-yr storm will become less common).

```{r, fig.height=4.8*3}

s <- 
  "/mnt/pers_disk/lev_and_prob_mosaics/mos_prob-b99perc_a1_staticregs.nc" %>% 
  read_ncdf() 

s[is.na(land)] <- NA

l_p <- 
  imap(seq(1, 3, by = 0.5) %>% as.character, function(wl, wl_i){
    
    ss <- 
      s %>% 
      slice(wl, wl_i) %>% 
      setNames("v") %>% 
      mutate(v = round(v, 2))
    
    d <- 
      ss %>% 
      pull() %>% 
      .[. > 1] %>%
      quantile(seq(0,1, by = 0.1), na.rm = T)
    
    tb_neg <- 
      ss %>% 
      as_tibble() %>% 
      mutate(wl = {{wl}}) %>% 
      filter(v <= 1) %>% 
      mutate(v = factor(-1))
    
    tb <-
      ss %>% 
      mutate(v = ifelse(v <= 1, NA, v)) %>% 
      cut(d) %>% 
      # mutate(v = as.integer(v)) %>%
      as_tibble() %>% 
      mutate(wl = {{wl}}) %>% 
      bind_rows(tb_neg)
    
    ggplot() +
      geom_raster(data = tb %>% filter(v != "-1"), aes(lon, lat, fill = v)) +
      geom_raster(data = tb %>% filter(v == "-1"), aes(lon, lat), fill = "grey70") +
      coord_equal(ylim = c(-60, 78), xlim = c(-160, 155)) +
      theme(axis.title = element_blank(),
            legend.position = "bottom") +

      scale_x_continuous(expand = c(0, 0)) +
      scale_y_continuous(expand = c(0, 0)) +

      guides(fill = guide_legend(nrow = 2, keywidth = 0.8, keyheight = 0.8, byrow = T)) +

      scale_fill_discrete_sequential("plasma",
                                     rev = T,
                                     na.value = "transparent",
                                     name = str_glue("Decile   ")) +
      labs(subtitle = str_glue("WL: {wl} degrees"))
    
  })

patchwork::wrap_plots(l_p[c(1,3,5)], ncol = 1)

```

---

The following figure shows the differences between v1 and v2 of the 100-yr storm precipitation amounts maps. Blue (red) represent areas where the new map depicts higher (lower) amounts than the old map. The map at the top, corresponding to a warming level of 0.5 degrees, shows "raw" precipitation differences. The middle and bottom map show differences in _change_ of precipitation (i.e. differences of deltas).

```{r}

dir_originals <- "/mnt/pers_disk/originals"

# if(!dir.exists(dir_originals)){
#   dir.create(dir_originals)
# }
# 
# "gsutil -m cp gs://wcdi_production/water_module/rcm_regcm_remo/global_RegCM_and_REMO_pr_100yrstorm.nc {dir_originals}/" %>%
#   str_glue() %>%
#   system()
# 
# "gsutil -m cp gs://wcdi_production/water_module/rcm_regcm_remo/global_RegCM_and_REMO_rp_100yrstorm.nc {dir_originals}/" %>%
#   str_glue() %>%
#   system()

l_s <- 
  map2(c("_pr_", "_rp_"), 
       c("_99perc_", "_prob-b99perc_"), 
       function(v_old, v_new){
         
         if(v_old == "_pr_"){
           slices_1 = c(1,3,6)
           slices_2 = c(1,3,6)
           wls = seq(0.5, 3, by = 0.5)
         } else {
           slices_1 = c(2,4,6)
           slices_2 = c(1,3,5)
           wls = seq(1, 3, by = 0.5)
         }
         
         s_old <- 
           dir_originals %>%
           list.files(full.names = T) %>% 
           str_subset(v_old) %>%
           read_ncdf(make_time = F) %>% 
           select(2) %>% 
           slice(time, slices_1)
         
         s_new <- 
           "/mnt/pers_disk/lev_and_prob_mosaics" %>% 
           list.files(full.names = T) %>% 
           str_subset(v_new) %>%
           str_subset("a1_staticregs") %>% 
           read_ncdf() %>% 
           st_set_dimensions("wl", values = wls) %>% 
           slice(wl, slices_2) %>% 
           setNames("v")
         
         s_old <- 
           s_old %>% 
           st_warp(s_new) %>% 
           st_set_dimensions(3, names = "wl") %>% 
           setNames("v") %>%
           mutate(v = set_units(v, NULL),
                  v = ifelse(v == -99999, NA, v))
         
         list(new = s_new, old = s_old)
    
  })


l_s <- 
  map(l_s, function(l){
    
    map(l, function(s){
      s[is.na(land)] <- NA
      return(s)
    })
    
  })

l_dif <- 
  map(l_s, function(l){
    
    dif <- l[[1]] - l[[2]]
    return(dif)
    
  })



# Precipitation amount

tb <- 
    l_dif[[1]] %>% 
    as_tibble()


tb <-
  tb %>% 
  mutate(v = raster::clamp(v, -200, 200))

tb %>% 
  ggplot(aes(lon, lat, fill = v)) +
  geom_raster() +
  facet_wrap(~wl, ncol = 1) +
  coord_equal(ylim = c(-60, 78), xlim = c(-160, 155)) +
  theme(axis.title = element_blank(),
        legend.position = "bottom") +
  
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  
  guides(fill = guide_coloursteps(barwidth = 20, 
                                  barheight = 0.5, 
                                  ticks = F, 
                                  show.limits = F,
                                  even.steps = F)) +
  
  colorspace::scale_fill_binned_diverging("blue-red3",
                                          na.value = "transparent",
                                          breaks = c(-200, -150, -90, -50, -10, 10, 50, 90, 150, 200),
                                          rev = T,
                                          name = "mm   ")




```

--- 

The following maps show differences in the _change_ of frequency of the 100-yr storm between v1 and v2 maps (i.e. all maps show differences of deltas). Blue (red) represent areas where the new map depicts higher (lower) frequency than the old map.

```{r}

tb <- 
    l_dif[[2]] %>% 
    as_tibble()

tb <-
  tb %>%
  mutate(v = raster::clamp(v, -4, 4))

tb %>% 
  ggplot(aes(lon, lat, fill = v)) +
  geom_raster() +
  facet_wrap(~wl, ncol = 1) +
  coord_equal(ylim = c(-60, 78), xlim = c(-160, 155)) +
  theme(axis.title = element_blank(),
        legend.position = "bottom") +
  
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  
  guides(fill = guide_coloursteps(barwidth = 20, 
                                  barheight = 0.5, 
                                  ticks = F, 
                                  show.limits = F,
                                  even.steps = F)) +
  
  colorspace::scale_fill_binned_diverging("blue-red3",
                                          na.value = "transparent",
                                          breaks = c(-4, -3, -2, -1, -0.5, 0.5, 1, 2, 3, 4),
                                          rev = T,
                                          name = "x (times)   ")
```


















