---
title: "Regional Frequency Analysis: Sliding Window Approach"
author: "Carlos Dobler -"
date: "`r Sys.Date()`"
output: html_document

---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = F,
                      message = F,
                      warning = F,
                      out.width="90%")


library(tidyverse)
library(lubridate)
library(stars)
library(furrr)
library(units)
library(tmap)

options(future.fork.enable = T)
plan(multicore)

dir_tmp <- "/mnt/pers_disk"

```


An alternative to using pre-defined regions to fit a regional GEV distribution is to use a (spatially) sliding window. This would be akin to RFA approaches based on a circular search area centered on a site of interest. Under these approaches, the neighborhood around a site is considered a "region". Because these "regions" vary only slightly between adjacent sites, distributions and derived statistics vary relatively smoothly across space. Moreover, a sliding window approach overcomes the challenge of accounting for shifts in pre-defined regions due to climate change.

This report shows results of calculating the precipitation and probability change of a 100 year storm using a sliding window method. The steps I followed were:

1. Calculate annual block maximas of precipitation for each model for each domain.
2. Slice each model by warming levels.
3. Pool (concatenate) block maximas from all models by warming level.
4. For each warming level, fit a regional GEV using a 5x5 sliding window as region, and obtain the 99th percentile and exceedance probability of the central grid cell.
5. Mosaic both variables from all domains for all warming levels.

.....  

Change in precipitation of a 100 year storm under 1.0 (left) and 3.0 (right) degrees of warming:

```{r}

s <-
  "/mnt/pers_disk/lev_and_prob_mosaics/mos_99perc_a2.nc" %>%
  read_ncdf()

land <-
  "/mnt/bucket_mine/misc_data/ne_110m_land/ne_110m_land.shp" %>%
  st_read(quiet = T) %>%
  mutate(a = 1) %>%
  select(a) %>%
  st_rasterize(s %>% slice(wl, 1) %>% mutate(b = NA) %>% select(b))

s <-
  s %>%
  split("wl") %>%
  select(2,6) %>%
  setNames(c("wl_1.0", "wl_3.0"))

s[is.na(land)] <- NA

s <- s %>% .[,1:1798,]

tmap_mode("view")

tm_shape(s) +
  tm_raster(palette = hcl.colors(20, "Zissou1", rev = T),
            breaks = c(-600, -1, 11, 24, 50, 999),
            labels = c("< -1", "-1 - 11", "12 - 24", "25 - 50", "> 50")) +
  tm_facets(sync = T,
            nrow = 1) +
  tm_view(set.view = 3)


```

.....  

Change in frequency of a 100 year storm under 1.0 (left) and 3.0 (right) degrees of warming:

```{r}

s <-
  "/mnt/pers_disk/lev_and_prob_mosaics/mos_prob-b99perc_a2.nc" %>%
  read_ncdf()

s <-
  s %>%
  split("wl") %>%
  select(1,5) %>%
  setNames(c("wl_1.0", "wl_3.0"))

s[is.na(land)] <- NA

s <- s %>% .[,1:1798,]

tm_shape(s) +
  tm_raster(palette = hcl.colors(20, "Zissou 1", rev = T),
            breaks = c(-600, 1, 2, 3, 4, 999),
            labels = c("< 1", "1 - 2", "2 - 3", "3 - 4", "> 4"),
            mid = 1.5) +
  tm_facets(sync = T,
            nrow = 1) +
  tm_view(set.view = 3)


```






















